CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes as ground truth for deep learning

### 2019-09-17 PaperRMFI2019/samplepaper.tex
more on calvis

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
\usepackage{amsmath}
\usepackage{amssymb}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\renewcommand\UrlFont{\color{blue}\rmfamily}

@@

\keywords{Human Shape Estimation \and CNN \and Another keyword.}
\end{abstract}
%
%
%
\textbf{Motivation} 
Predicting 3D human body measurements from images is crucial in several 
scenarios 
like virtual try-on, animating, ergonomics, computational forensics and 
even health and mortality estimation. Researches had 
named this 
measurements body intuitive controls (\cite{Allen.2003}), biometric 
measurements (\cite{Sigal.2008}), body dimensions 
(\cite{DBLP:conf/bmvc/ChenRC11}, European standard EN 13402), semantical 
parameters (\cite{Yang.2014}), 
traditional anthropometric 
measurements 
(\cite{Wuhrer2011}) or only ``shape" as in human shape estimation 
(\cite{Guan.2013}, \cite{Bogo:ECCV:2016}, \cite{Loper.2015}, 
\cite{Dibra.2016a}, \cite{Pishchulin.2017}). In contrast, we assume a more 
anthropometric approach 
motivated by comprehensive compendiums like Panero and Zelnik, 1979 
\cite{panero1979human}.
Throughout this paper the term human body 
dimensions 
will be used to refer to the above measurements.

@@

Problem statement: given a 3D human body mesh $\mathcal{M}$ we look for a 
method capable to automatically output chest, waist and pelvis circumference.

%------------------------------------------------------------------------
\section{Approach}\label{sec:approach}

In this work we synthesize 3D human meshes using the SMPL
model \cite{Loper.2015}. The generative nature of our approach establishes this 
model as starting point (and not the 3D mesh). Nevertheless, our method is 
flexible enough to begin the pipeline with a 3D mesh. In that case, an SMPL 
model can be fitted to it, using the method of (rh) for example.
This model is at its top level a skinned articulated 
model, i.e., 
consists of a 
surface mesh that mimics the skin and a skeleton related to that mesh. It is 
defined by a mean 
template shape represented by a vector of $N$ concatenated vertices 
$\mathbf{\bar{T}}$ in the zero pose, $\vec{\theta}^*$. In order 
to 
synthesize a 
new human mesh one has to deform the provided template mesh by 
setting shape parameters $\vec{\beta}$ and pose parameters $\vec{\theta}$. The 
model provides learned parameters
\begin{equation} \label{eq:smpl_params}
\Phi = \{\mathbf{\bar{T}}, \mathcal{W}, \mathcal{S}, \mathcal{J}, 
\mathcal{P}\}
\end{equation}
As mentioned above $\mathbf{\bar{T}}$ is the mean template shape. The weight 
matrix $\mathcal{W}$ represents how much the rotation of skeleton parts affects 
the vertices. In addition, the matrices $\mathcal{S}$ and $\mathcal{P}$ define 
linear functions that are used to deform $\mathbf{\bar{T}}$ and the matrix 
$\mathcal{J}$ predicts skeleton rest joint locations from vertices in the rest 
pose. We held fix these parameters during the synthesis.

@@

@@

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\renewcommand\UrlFont{\color{blue}\rmfamily}

@@

\keywords{Human Shape Estimation \and CNN \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction}\label{sec:intro}
\textbf{Motivation} 
Predicting 3D human body measurements from images is crucial in several 
scenarios 
like virtual try-on, animating, ergonomics, computational forensics and 
even health and mortality estimation. Researches had 
named this 
measurements body intuitive controls (\cite{Allen.2003}), biometric 
measurements (\cite{Sigal.2008}), body dimensions 
(\cite{DBLP:conf/bmvc/ChenRC11}, European standard EN 13402), semantical 
parameters (\cite{Yang.2014}), 
traditional anthropometric 
measurements 
(\cite{Wuhrer2011}) or only ``shape" as in human shape estimation 
(\cite{Guan.2013}, \cite{Bogo:ECCV:2016}, \cite{Loper.2015}, 
\cite{Dibra.2016a}, \cite{Pishchulin.2017}). In contrast, we assume a more 
anthropometric approach 
motivated by comprehensive compendiums like Panero and Zelnik, 1979 
\cite{panero1979human}.
Throughout this paper the term human body 
dimensions 
will be used to refer to the above measurements.

@@

\textbf{Problem statement:} given a 3D human body mesh $\mathcal{M}$ we look 
for a 
method capable to automatically output chest, waist and pelvis circumference.

\section{Related Work}\label{sec:rel_work}

%------------------------------------------------------------------------
\section{Approach}\label{sec:approach}
To calculate the actual human body dimensions, we first need to formalize the 
notion of chest, waist and pelvis of a 3D human body mesh. The starting point 
is the model capable to represent that mesh, which we discuss in subsection 
\ref{subsec:smpl_model}. Next, we define a mesh signature over that mesh in 
subsection \ref{eq:mesh_signature} that allow us to formalize the human body 
dimensions.
Our strategy segmentates consist of segmentating the mesh torso in three 
regions, namely chest, waist and pelvic regions. We then apply the formal 
definitions and search for extrema and identify points in those regions that we 
can interpret as the wanted body dimensions.

\subsection{SMPL Model of the Human Body}\label{subsec:smpl_model}
In this work we synthesize 3D human meshes using the SMPL
model \cite{Loper.2015}. The generative nature of our approach establishes this 
model as starting point (and not the 3D mesh). Nevertheless, our method is 
flexible enough to begin the pipeline with a 3D mesh. In that case, an SMPL 
model can be fitted to it, using the method of (rh) for example.
This model is at its top level a skinned articulated 
model, i.e., 
consists of a 
surface mesh that mimics the skin and a skeleton related to that mesh. It is 
defined by a mean 
template shape represented by a vector of $N$ concatenated vertices 
$\mathbf{\bar{T}}$ in the zero pose, $\vec{\theta}^*$. In order 
to 
synthesize a 
new human mesh one has to deform the provided template mesh by 
setting shape parameters $\vec{\beta}$ and pose parameters $\vec{\theta}$. The 
model provides learned parameters
\begin{equation} \label{eq:smpl_params}
\Phi = \{\mathbf{\bar{T}}, \mathcal{W}, \mathcal{S}, \mathcal{J}, 
\mathcal{P}\}
\end{equation}
As mentioned above $\mathbf{\bar{T}}$ is the mean template shape. The weight 
matrix $\mathcal{W}$ represents how much the rotation of skeleton parts affects 
the vertices. In addition, the matrices $\mathcal{S}$ and $\mathcal{P}$ define 
linear functions that are used to deform $\mathbf{\bar{T}}$ and the matrix 
$\mathcal{J}$ predicts skeleton rest joint locations from vertices in the rest 
pose. We held fix these parameters during the synthesis.

@@

\subsection{Chest, Waist and Pelvis Regions 
Segmentation}\label{subsec:three_regions}

\subsubsection{Axilla recognition}\label{subsec:armpit_recog}
As mentioned above, we would like to measure the chest circumference below the 
armpit, also known as axilla. This raises the question whether we can recognize 
the axilla based on 
the mesh signature. We could then establish the location index as lower bound 
for the search of chest location. There is consensus regarding the axilla 
definition as the area of the human body underneath the shoulder joint 
(\cite{MeSH.axilla}, \cite{FMA.axilla}, \cite{TA.axilla}). Therefore, we can 
use the shoulder joint as a hint to locate the axilla and establish a lower 
bound for the chest region. We then can combine this idea with the defintion 
used in the garment industry, namely the maximum horizontal girth under the 
armpits (axillae) (rh).
Let us consider the left shoulder joint on in our skeleton model $J_{ls}$. We 
cast a ray in direction $d$ outside the mesh. Since the joint $J_{ls}$ lies 
inside the mesh, the ray intersects its surface at point $p_{axilla}$ (see 
figure x).
With this construction we can define the chest region $\mathcal{CR}$ as the 
region where the slice index is bigger that the index of a slice with 
y-coordinate is equal or greater than the axilla $y$-coordinate 
$y_{p_{axilla}}$.

\begin{align}
\mathcal{CR} = \mathcal{BL}(j), \, j \in \{j_t, j_t+m, \cdots, j_b\}
\end{align}

Similarly, the waist region is

\begin{align}
\mathcal{WR} = \mathcal{BL}(j), \, j \in \{j_t, j_t+m, \cdots, j_b\}
\end{align}

Finally, the pelvic region is

\begin{align}
\mathcal{PR} = \mathcal{BL}(j), \, j \in \{j_t, j_t+m, \cdots, j_b\}
\end{align}

\subsection{Recognition by interval search}\label{subsec:interval_search}

@@

\subsection{Qualitative evaluation}\label{subsec:qualitative_eval}
Here we evaluate our method qualitatively with method one and method two.
\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{subject_6_with_cutting_planes_on_joints.eps}
	\end{center}
	\caption{Qualitative evaluation for male and female meshes. We evaluate our 
	method with method one and method two.}
	\label{fig:qualitative_eval}
\end{figure}

Figure \ref{fig:qualitative_eval} shows the $\mathcal{MS}$ of two meshes (male 
and female) for 
$m=0,001$ and plane $\boldsymbol{\pi}$ parallel to the floor (with normal $(0, 
1,0)$) using the library trimesh \cite{trimesh}. Note that the function as 
defined by equation \ref{eq:mesh_signature} is bounded and not continuous. It 
resembles a rotated silhouette of the human body and exhibits several 
\textit{extrema}.

\begin{figure}[h]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:qualitative_eval}
\end{figure}

\subsection{Learning convergence evaluation}\label{subsec:learn_conv}
This experiment demonstrates the plausibility of using calvis to obtain data 
that can be used as a ground truth for machine learning (in general) and 
specifically as supervising signal for ANN.

Figure \ref{fig:calvis_net} shows the our prototypical ANN calvis net where the 
input is a synthetic image from a 3D mesh and the output are the three body 
dimensions.

Figure \ref{fig:learn_conv} shows that after epoch number 5 the estimation 
error decreases and the learning converges.

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:qualitative_eval}
\end{figure}

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:qualitative_eval}
\end{figure}
